prep_annotating

### 2022-06-10 code/prep_annotate.Rmd
process data for annotating flower referents

---
title: "aa_mic_submission_analyses"
author: "Jessica Mankewitz"
date: "9/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This notebook reads in the preprocessed and joined data, runs the basic processing/summarizing code, and generates figures and numbers for the 2021 MIC workshop submission.

```{r}
library(here)
library(tidyverse)

#for plotting
source(here("code/theme_specs.R"))

@@

# Summary Stats
## Game counts
```{r}
d.games.final %>% 
  group_by(-game_complete, condition, chatEnabled) %>%
  count()
```

## Game lengths
```{r}
d.games.final %>% 
filter(game_complete) %>%
  group_by(condition, chatEnabled) %>%
  mutate(gameLength_minutes = gameLength/60) %>%
  summarize_at("gameLength_minutes", funs(min, max, mean))
```

## Bonus Payments
Max bonus payment
```{r}
max(d.players$bonus, na.rm=T)
```
## Hourly compensation 
```{r}
#I just pulled these averages from prolific
(35.81*16 + 25.75 * 47 + 26.38 * 47 + 24.76*18 + 10.17* 18 + 12.64 * 18 + 11.79 * 68 + 11.24*66)/(16+47+47+18+18+18+68+66)

@@

# Summarizing Data

We'll pull out the summary statistics with 3 views: 
* by player (points earned, number of utterances, number of words)
  ** selection metrics: select hidden, non-top flower
* by round (group points earned, max possible points earned in round, number of words exchanged by group)
* by game (total points, max possible points in game)

```{r}
#player language per round
d.player_word_counts <- d.raw_chat.final %>% 
  filter(type == "message") %>%
  mutate(text = gsub('[[:punct:] ]+',' ',text),
         text = str_squish(text),
         utt_length_chars = str_length(text), 
         utt_length_words = str_count(text, "\\W+") + 1) %>% 
  group_by(gameId, blockNum, trialNum, repNum, playerId, numPlayers) %>%
  summarize(text = paste0(text, collapse = ' - '),
            total_num_words = sum(utt_length_words),
            total_num_chars = sum(utt_length_chars))

# player points per round
d.by_player_metrics <- d.rounds.final  %>% 
  left_join(d.player_word_counts) %>%
  left_join(d.games.conditions) %>% 
  replace_na(list(total_num_words = 0, total_num_chars = 0))

#add deference measure and flower ranking
deference_rank <-  d.by_player_metrics %>%  
  full_join(d.contexts %>% 
              select(roundId, blinded, label, utility)) %>%
  filter(blinded != playerId) %>%
  group_by(playerId, roundId) %>%
  arrange(-utility) %>%
  mutate(selectionRank = c(0:3)) %>%
  filter(playerResponse == label) %>%
  mutate(deference = selectionRank != 1) %>%
  select(roundId, playerId,deference, selectionRank)

d.by_player_metrics <- d.by_player_metrics %>% 
  left_join(d.contexts %>% select(roundId, playerResponse = label, 
                                  selectedHidden = blinded,
                                  selectionUtility = utility)) %>%
  mutate(selectedHidden = selectedHidden == playerId) %>%
  left_join(deference_rank)
```

Calculate the best possible per round

```{r max_round}
d.max_round <- d.contexts %>%
  group_by(gameId, repNum, blockNum, trialNum) %>%
  #top 3 flowers
  slice_max(utility, n = 3) %>%
  summarise(roundMax = sum(utility))
```

```{r}
d.by_round_metrics <- d.by_player_metrics  %>% 
  group_by(gameId, trialNum, roundId) %>%
  summarize(roundUtility = sum(playerUtility, na.rm = T),
            numWords = sum(total_num_words, na.rm = T),
            numChars = sum(total_num_chars, na.rm = T),
            numFlowers = n_distinct(playerResponse),
            roundLength = max(time_sec)) %>%
  left_join(d.max_round) %>%
  mutate(pointsAdjusted = roundUtility/roundMax) %>% 
  left_join(d.games.conditions)
```
```{r}
d.by_game_metrics <- d.by_round_metrics  %>% 
  group_by(gameId) %>%
  summarize(gameUtility = sum(roundUtility, na.rm = T),
            numWords = sum(numWords, na.rm = T),
            numChars = sum(numChars, na.rm = T),
            maxPossPoints = sum(roundMax)) %>%
  mutate(pointsAdjusted = gameUtility/maxPossPoints) %>% 
  left_join(d.games.conditions) %>%
  left_join(d.games.final %>% select(gameId, gameLength))
```

# Merge in RL Experiments

```{r read-rl}
d.rl_metrics <- read_csv(here("data/processed_data/RL_experiment_results.csv"))

#get in line with player summary statistics?

d.rl.player_metrics <- d.rl_metrics %>% 
  mutate(gameId = paste(condition_raw, languageConditionRaw, trial, sep = "-"), 
         playerId = paste0(trial, "-",agent)) %>%
  select(playerId, languageCondition, condition, avgPlayerPoints = reward) %>% 
  mutate(rl_agent = "Agent")

# Fig 1 & 2
## RL Point Comp

### All Players
```{r}
d.by_player_metrics %>%
  group_by(languageCondition, condition, playerId) %>%
  summarize(avgPlayerPoints = mean(playerUtility)) %>% 
  mutate(rl_agent = "Human") %>%
  rbind(d.rl.player_metrics) %>%
  ggplot(aes(x = condition, y = avgPlayerPoints, color = rl_agent)) + 
  geom_point(position=position_jitterdodge(dodge.width=.5, jitter.width = .1), alpha = .25) +
  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.5)) + 
  scale_color_brewer(palette="Dark2")+ theme_jmank() + 
  facet_wrap(~languageCondition)+ 
  theme(legend.position = "bottom", axis.title.x=element_blank()) +
  labs(y = "Average Player Points", color = "Player Type")
```
### Avg Points over game
Do the same but by average points per round across players

```{r}
d.rl.game_metrics <- d.rl_metrics %>% 
  mutate(gameId = paste(condition_raw, languageConditionRaw, trial, sep = "-"), 
         playerId = paste0(trial, "-",agent)) %>%
  group_by(gameId, languageCondition, condition) %>% 
  summarize(avgRoundPoints = sum(reward)) %>%
  mutate(rl_agent = "Agent")

d.rl.game_metrics
```

{} to match other plot, facet by utility condition, color by agent type?
```{r}
rl_comparison <- d.by_round_metrics %>%
  group_by(languageCondition, condition, gameId) %>%
  summarize(avgRoundPoints = mean(roundUtility)) %>% 
  mutate(rl_agent = "Human") %>%
  rbind(d.rl.game_metrics) %>%
  ggplot(aes(x = languageCondition, y = avgRoundPoints, color = condition)) + 
  geom_point(position=position_jitterdodge(dodge.width=.5, jitter.width = .1), alpha = .25) +
  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.5)) + 
  scale_color_bay()+ theme_jmank() + 
 # labs(y = "Average Points", color = "Player Type") +
  labs(y = "Average Points per Trial", x = "Language Condition", color = "Utility Condition") +
  facet_wrap(~rl_agent)+ 
  theme(legend.position = "bottom",
        #legend.box="vertical",
        legend.margin=margin(),
        text = element_text(size=15),
        axis.text = element_text(size = 12),
        legend.title=element_text(size=12),
        legend.text=element_text(size=12),
        strip.text = element_text(size=15))

rl_comparison
```
```{r}
d.by_round_metrics %>%
  group_by(languageCondition, condition, gameId) %>%
  summarize(avgRoundPoints = mean(roundUtility)) %>% 
  mutate(rl_agent = "Human") %>%
  rbind(d.rl.game_metrics) %>%
  ggplot(aes(x = languageCondition, y = avgRoundPoints, color = condition)) + 
  geom_point(position=position_jitterdodge(dodge.width=.5, jitter.width = .1), alpha = .25) +
  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.5)) + 
  scale_color_brewer(palette="Dark2")+ theme_jmank() + 
 # labs(y = "Average Points", color = "Player Type") +
  labs(y = "Average Points per Round", color = "Player Type", x = "Language Condition") +
  facet_wrap(~rl_agent)+ 
  theme(legend.position = "bottom")
```

## Points x Time x Num. Words
```{r}
lang_by_points <- d.by_game_metrics %>% 
  ggplot(aes(x = gameLength/60, y = gameUtility, shape = languageCondition, color = log(numWords))) + 
  geom_point(alpha = .8, size = 4) +
  geom_smooth(aes(group = condition), method=lm, formula=y~log(x),color = "black")+
  facet_grid(cols = vars(condition)) + 
  labs(x = "Game length (in Minutes)", 
       y = "Points Earned by Group",
       color = "Log N. Words",
       shape = "") + 
  theme_jmank() +  scale_color_steps(low = "#edd746", high = "#6B9555") + 
  theme(legend.position = "bottom",
        #legend.box="vertical",
        legend.margin=margin(),
        text = element_text(size=15),
        axis.text = element_text(size = 12),
        legend.title=element_text(size=12),
        legend.text=element_text(size=12),
        strip.text = element_text(size=15))
lang_by_points
```

### Rewards Blurb

```{r}
d.by_game_metrics %>% mutate(lang_used = numWords > 0) %>%
  group_by(languageCondition, lang_used) %>% count()

max(d.by_game_metrics$numWords)
```
# Language Reduction

```{r}
# number of words exchanged decreases over time
lang_reduction <- d.by_round_metrics %>%
  filter(languageCondition == "Lang") %>%
  mutate(utility_cond_short = ifelse(condition == "Shared Utilities", "Shared", "Individual")) %>%
  ggplot(aes(x = trialNum, y = numWords, color = utility_cond_short)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  #geom_smooth(method=lm, formula=y~log(x))+
   labs(x = "Trial Number", y = "N. Words", color = "Utility Condition")+ 
  scale_color_bay() + theme_jmank() +theme(legend.position = "bottom") + 
   theme(legend.position = "bottom",
        text = element_text(size=15),
        axis.text = element_text(size = 12),
        legend.title=element_text(size=12),
        legend.text=element_text(size=12),
        strip.text = element_text(size=15))
lang_reduction

utterances <- d.raw_chat.final %>% 
  left_join(d.games.conditions) %>%
  filter(type == "message") %>%
  select(gameId, trialNum, playerId, text) %>%
  #this drops the NAs, we want to keep instances where there was a selection but no message ugh
  mutate(text = gsub('[[\']+]|[[:punct:] ]+',' ',text),
         text = str_squish(text),
         utt_length_chars = str_length(text), 
         utt_length_words = str_count(text, "\\W+")+1) %>%
  group_by(gameId, trialNum, playerId) %>%
  summarize(utt_length_chars = list(utt_length_chars),
            utt_length_words = list(utt_length_words))

# normal MLU also decreases
utt_mlu <- utterances %>% 
  unnest(c(utt_length_chars, utt_length_words)) %>%
  left_join(d.games.conditions) %>%
  group_by(trialNum, gameId) %>%
  mutate(mlu = mean(utt_length_words)) %>% 
  ggplot(aes(x = trialNum, y = mlu, color = condition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  labs(x = "Trial Number", y = "Mean Length of Utterance", color = "Utility Condition")+ 
  scale_color_bay() + theme_jmank() +theme(legend.position = "bottom")

# MLU including trials with 0 "length" utterances
expand_utt <- d.raw_chat.final %>%
  select(gameId, trialNum, playerId) %>% 
  distinct() %>%
  full_join(utterances)

no_lang_index <- sapply(expand_utt$utt_length_chars, is.null)
expand_utt$utt_length_chars[no_lang_index] <- 0
expand_utt$utt_length_words[no_lang_index] <- 0

utt_reduction <- expand_utt %>% 
  unnest(cols = c(utt_length_chars, utt_length_words)) %>%
  left_join(d.games.conditions) %>%
  filter(languageCondition == "Lang", game_complete) %>%
  group_by(gameId, trialNum, languageCondition, condition) %>%
  summarize(mean_utt_length = mean(utt_length_words))

utt_reduction %>% ggplot(aes(x = trialNum+1, y = mean_utt_length, color = condition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  labs(x = "Trial Number", y = "Average N. Words per Utterance", color = "Utility Condition")+ 
  scale_color_bay() + theme_jmank() +theme(legend.position = "bottom")

#points increase (slightly) over time
d.by_round_metrics %>%
  filter(languageCondition == "Lang") %>%
  ggplot(aes(x = trialNum, y = pointsAdjusted, color = condition)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method=lm)+
  scale_color_bay() + theme_jmank()

```
Point x Trial correlation
```{r}
library(psych)
corel <- d.by_round_metrics %>%
  filter(languageCondition == "Lang") %>% ungroup() %>%
  select(trialNum, pointsAdjusted, numWords)
corr.test(corel$trialNum, corel$pointsAdjusted)
```

# Final Figs

```{r}
library(patchwork)
utt_mlu
lang_by_points
rl_comparison
lang_reduction

lang_by_points + lang_reduction +plot_annotation(tag_levels = 'A')#, labels = c('A', 'B'), label_size = 12, aligh = "hv")

ggsave(here(fig_path, "points_earned_reduction_joined.png"), height = 4, width = 11)

rl_comparison
ggsave(here(fig_path, "rl_comparison.png"), height = 3, width = 5)
```

# Number of questions
Adapted from Georgia's report

```{r}
question_words <- c("who",  "what",  "where",  "why",  "how")
d_messages <- d.raw_chat.final %>% 
  filter(type=="message") %>%
  filter(!is.na(text))

d_messages <- d_messages %>% 
  mutate(question_mark = ifelse(str_detect(text, "\\?"),TRUE,FALSE),
         question_word = grepl(paste(question_words, collapse="|"), tolower(text)),
         any_question = ifelse((question_mark | question_word), TRUE, FALSE))

d_messaged_summarized <- d_messages %>% group_by(gameId, trialNum, condition) %>% 
  summarize(total_utt = n(), 
            n_question_marks = sum(question_mark), 
            n_question_words = sum(question_word),
            n_questions = sum(any_question),
            proportion_q = n_question_marks/total_utt)

d_messaged_summarized %>%
  ggplot(aes(x = trialNum, y = proportion_q, color = condition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  #geom_smooth(method=lm, formula=y~log(x))+
   labs(x = "Trial Number", y = "Prop of Questions", color = "Utility Condition")+ 
  scale_color_bay() + theme_jmank() + theme(legend.position = "none") + ggtitle("All Trials")

>>>>

---
title: "prep_annotating"
author: "vboyce"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)

library(here)
library(tidyverse)

#for plotting

@@

# Want to find a "best" game to start with

```{r}
d.rounds.final %>% filter(game_complete) %>% 
  group_by(gameId) %>% summarise(utils=sum(playerUtility)) %>% arrange(desc(utils))

# let's look at gameId == QHkfso2yMgbbziE7r first
```

```{r}

good_rounds <- d.rounds.final %>% filter(gameId=="QHkfso2yMgbbziE7r") %>% 
  select(playerId,playerResponse,roundId,gameId) %>% 
  separate(playerResponse, into=c("color","pResp"))

good_contexts <- d.contexts %>% filter(gameId=="QHkfso2yMgbbziE7r") %>% 
  select(roundId, gameId, label, utility) %>% 
  separate(label, into=c("color", "num")) %>% 
  mutate(labelutil=str_c(num,":",utility)) %>% 
  group_by(roundId, gameId, color) %>% 
  arrange(desc(utility)) %>% 
  summarize(flowers=str_c(labelutil,collapse=", "))

good_chat <- d.raw_chat.final %>% filter(gameId=="QHkfso2yMgbbziE7r") %>% 
  left_join(d.players %>% select(playerId, name)) %>% 
  rename(roundId=roundID) %>% 
  left_join(good_contexts) %>% 
  left_join(good_rounds) %>% 
  mutate(rowid=row_number()) %>% 
  select(rowid,trialNum,name, pResp, text, flowers, color) %>% 
  filter(!is.na(text)) %>% 
  write_tsv("test.tsv")

@@
